Search for Spatially Extended \fermi-LAT Sources Using Two Years of Data

### 2012-05-17 extended_catalog/paper/extended_search.tex
upew draft v13

\subsection{Point-like Source Simulations Over a Uniform Background}

@@

\label{monte_carlo_validation}

@@

\subsection{\newtext{Point-like Source Simulations Over a Structured Background}}

We performed a second set of simulation to show that the theoretical distribution
for \tsext is still preserved when the point-like sources are present over
a highly-structured diffuse background.
Our simulation setup was the same as above except that the sources were
simulated on top of and analyzed assuming the presence of the standard
Galactic diffuse and isotropic background models used in 2FGL.  In our
simulations, we selected our sources to have random positions on the sky
such that they were within 5\degree of the Galactic plane. This probes the 
brightest and most strongly contrasting areas of the Galactic background.

To limit the number of tests, we selected only one flux
level for each of the four spectral indices and we performed
this test only in the 1 \gev to 100 \gev energy range. 
As described below, The fluxes were selected so that $\ts\sim50$. We do not
expect to be able to spatially resolve sources much dimmer than this
and the results for sources much brighter are less likely to be affected by the 
structured background.

Because the Galactic diffuse emission is highly structure with
strong gradients, the point-source
sensitivity can vary significantly across the Galactic plane.
To account for this, we scaled the flux (for a given spectral index)
so that the source always has approximately the same signal-to-noise ratio:
\begin{equation}
  \label{scale_flux_by_background}
  F(\vec{x}) = F(\text{GC}) \times \left(
  \frac{B(\vec{x})}{B(\text{GC})}\right)^{1/2}.
\end{equation}
Here, $\vec{x}$ is the position of the simulated source, $F$ is the integral
flux of the source from 100 \mev to 100 \gev, $F(\text{GC})$
is the same quantity if the source was at the Galactic center, $B$
is the integral of the Galactic diffuse and isotropic emission
from 1 \gev to 100 \gev at the position of the source, and $B(\text{GC})$ is the same quantity
if the source was at the Galactic center.  For the four spectral models,
Table~\ref{ts_ext_num_sims} lists $F(\text{GC})$ and the average value of \ts.

@@

We used a simulation setup similar to that described in
Section~\ref{monte_carlo_validation}, but the sources were simulated on
top of and analyzed assuming the presence of the standard Galactic diffuse and
isotropic background model used in 2FGL.  
We simulated the gamma-ray emission in the 1 \gev to 100 \gev
energy range for the 2-year interval of the 2FGL catalog.
In the simulations, 
we did not include the finite energy resolution of the LAT
to isolate any effects due to changing the assumed spatial model. 
The fitting code we use also ignores this energy dispersion and the
potential bias introduced by this will be discussed in an upcoming paper
by the LAT collaboration \citep{lat_on_orbit_psf}.
In total, we performed 985 independent simulations.

@@

Figure~\ref{ts_comparison_w44sim}a 
shows that
the significance of W44 in the simulations is very great ($\ts\sim3500$) 
for a model with a point-like source hypothesis.
Figure~\ref{ts_comparison_w44sim}b shows that 
the significance of the spatial extension is also very great
($\tsext\sim250$).  
On average \tsext is somewhat larger when fitting
the sources with more accurate spatial models.  This shows that
assuming an incorrect spatial model will cause the source's
significance to be underestimated.  Figure~\ref{ts_comparison_w44sim}c
shows that the sources were fit better when assuming an elliptical
disk spatial model compared to a uniform disk spatial model
($\ts_\text{elliptical\ disk}-\ts_\text{disk}\sim30$).  Finally,
Figure~\ref{ts_comparison_w44sim}d shows that the sources were
fit somewhat better assuming an elliptical ring spatial model
compared to an elliptical disk spatial model ($\ts_\text{elliptical\
ring}-\ts_\text{elliptical\ disk}\sim9$). This shows that the LAT has
some additional power to resolve substructure in bright extended sources.

Figure~\ref{bias_w44sim}a and Figure~\ref{bias_w44sim}b clearly show that
no significant bias is introduced by modeling the source as extended
but with an inaccurate spatial model, while a point-like source modeling
results in a $\sim10\%$ and $\sim5\%$ bias in the fit flux and index,
respectively.  
Furthermore, Figure~\ref{bias_w44sim}c shows that the \rsixeight estimate of
the extension size is very mildly biased ($\sim10\%$) toward higher values
when inaccurate spatial models are used, and thus represent a reasonable
measurement of the true 68\% containment radius for the source.
For the elliptical spatial models, \rsixeight is computed by numeric integration.

>>>>

\subsection{Point-like Source Simulations Over a Uniform Background}
\label{monte_carlo_validation}

@@

@@

\subsection{\newtext{Point-like Source Simulations Over a Structured Background}}
\label{validation_over_plane}

We performed a second set of simulations to show that the theoretical distribution
for \tsext is still preserved when the point-like sources are present over
a highly-structured diffuse background.
Our simulation setup was the same as above except that the sources were
simulated on top of and analyzed assuming the presence of the standard
Galactic diffuse and isotropic background models used in 2FGL.  In our
simulations, we selected our sources to have random positions on the sky
such that they were within 5\degree of the Galactic plane. This probes the 
brightest and most strongly contrasting areas of the Galactic background.

To limit the number of tests, we selected only one flux
level for each of the four spectral indices and we performed
this test only in the 1 \gev to 100 \gev energy range. 
As described below, the fluxes were selected so that $\ts\sim50$. We do not
expect to be able to spatially resolve sources much dimmer than this
and the results for sources much brighter are less likely to be affected by the 
structured background.

Because the Galactic diffuse emission is highly structured with
strong gradients, the point-source
sensitivity can vary significantly across the Galactic plane.
To account for this, we scaled the flux (for a given spectral index)
so that the source always has approximately the same signal-to-noise ratio:
\begin{equation}
  \label{scale_flux_by_background}
  F(\vec{x}) = F(\text{GC}) \times \left(
  \frac{B(\vec{x})}{B(\text{GC})}\right)^{1/2}.
\end{equation}
Here, $\vec{x}$ is the position of the simulated source, $F$ is the integral
flux of the source from 100 \mev to 100 \gev, $F(\text{GC})$
is the same quantity if the source was at the Galactic center, $B$
is the integral of the Galactic diffuse and isotropic emission
from 1 \gev to 100 \gev at the position of the source, and $B(\text{GC})$ is the same quantity
if the source was at the Galactic center.  For the four spectral models,
Table~\ref{ts_ext_num_sims} lists $F(\text{GC})$ and the average value of \ts.

@@

We used a simulation setup similar to that described in
Section~\ref{validation_over_plane}, but the simulations
were over the 2-year interval of the 2FGL catalog.
In the simulations, 
we did not include the finite energy resolution of the LAT
to isolate any effects due to changing the assumed spatial model. 
The fitting code we use also ignores this energy dispersion and the
potential bias introduced by this will be discussed in an upcoming paper
by the LAT collaboration \citep{lat_on_orbit_psf}.
In total, we performed 985 independent simulations.

@@

Figure~\ref{ts_comparison_w44sim}a 
shows that
the significance of W44 in the simulations is very large ($\ts\sim3500$) 
for a model with a point-like source hypothesis.
Figure~\ref{ts_comparison_w44sim}b shows that 
the significance of the spatial extension is also large
($\tsext\sim250$).  
On average \tsext is somewhat larger when fitting
the sources with more accurate spatial models.  This shows that
assuming an incorrect spatial model will cause the source's
significance to be underestimated.  Figure~\ref{ts_comparison_w44sim}c
shows that the sources were fit better when assuming an elliptical
disk spatial model compared to a uniform disk spatial model
($\ts_\text{elliptical\ disk}-\ts_\text{disk}\sim30$).  Finally,
Figure~\ref{ts_comparison_w44sim}d shows that the sources were
fit somewhat better assuming an elliptical ring spatial model
compared to an elliptical disk spatial model ($\ts_\text{elliptical\
ring}-\ts_\text{elliptical\ disk}\sim9$). This shows that the LAT has
some additional power to resolve substructure in bright extended sources.

Figure~\ref{bias_w44sim}a and Figure~\ref{bias_w44sim}b clearly show that
no significant bias is introduced by modeling the source as extended
but with an inaccurate spatial model, while a point-like source modeling
results in a $\sim10\%$ and $\sim0.125$ bias in the fit flux and index,
respectively.  
Furthermore, Figure~\ref{bias_w44sim}c shows that the \rsixeight estimate of
the extension size is very mildly biased ($\sim10\%$) toward higher values
when inaccurate spatial models are used, and thus represents a reasonable
measurement of the true 68\% containment radius for the source.
For the elliptical spatial models, \rsixeight is computed by numeric integration.
